Simplified-spatial-availability-example

### 2022-05-31 Simplified-spatial-availability-example.Rmd
Finished converting to ZT208 resolutions.
Am working on how to shift car trips to transit (in a way that's flexible and will allow us to easily toggle the modal shift scenarios (i.e., instead of 100% transit, maybe 80%/10/10 transit/bike/walk))

Import the trip data:
```{r data-import-filter}
OD_v <- read_excel("data-raw/EDM2018VIAJES.xlsx")
```

@@

Calculate the duration of each trip
```{r}
OD_v <- OD_v %>%
  mutate(TIEMPO_VIAJES =
           (as.numeric(OD_v$VDESHORAFIN) - 
              as.numeric(OD_v$VORIHORAINI))) %>%
  mutate(TIEMPO_VIAJES = ifelse(TIEMPO_VIAJES == 0, 0.1, TIEMPO_VIAJES)) # set all 0min travel times to 0.1 min
summary(OD_v$TIEMPO_VIAJES)
```
Checking the trip data - it appears each OD is a row, i.e., chained trips are multiple rows. We want to extract every home origin (VORI = 1) to every work destination (VDEST =2) trip as a single row... so let's do this in multiple steps.

First, the direct home to work trips:
```{r}
# not chained home to work trips - direct
OD_v_HW <- OD_v %>% filter(VORI == "1" & VDES == "2") #only filter in origin casa trips and destination trabajo
```

Next, let's identify the chained trips and then consolidate them. We only care about trip O/D, travel mode, travel time, and 
```{r}
t <- OD_v %>% group_by(ID_HOGAR,ID_IND) %>% 
  filter((VORI == "1" & VDES != "2" & VDES != "1") | (VDES == "2" & VORI != "1"& VORI != "2")) %>% #here the origin NEEDS to be from home, the destination can be anywhere OR the origin can be from anywhere but the destination NEEDS to be work. 
  mutate(chained = lag(VDES) - VORI)

chained2 <- (ifelse(t$chained[-1]==0 |t$chained==0, 0, NA)) %>% as.data.frame()

t <- cbind(t,chained2) %>% dplyr::rename("chained2" = ".")

t<-t %>% filter(chained2==0) # these are all the CHAINED trips - only considering 1 location before work.  
t
```

Now... let's only include chained trips with the same mode! 
```{r}
t <- t %>% group_by(ID_HOGAR, ID_IND) %>% mutate(trip = ifelse(((ID_VIAJE - lag(ID_VIAJE)) == 1)
                                                               & ((MODO_PRIORITARIO - lag(MODO_PRIORITARIO))== 0), 
                                                               MODO_PRIORITARIO, 0))

t %>% filter(trip == 0) %>% nrow() / t %>% filter(chained == 0) %>% nrow()
t %>% filter(trip != 0 & !is.na(trip)) %>% nrow() / t %>% filter(chained == 0) %>% nrow()
```
so ~27% of chained trips are not made using the same mode. THe rest of them are (`r t %>% filter(trip != 0 & !is.na(trip)) %>% nrow() ` trips are chained (1 intermediate destination) and use the same mode).

@@

Let's see if this sample increases the number of ZTs spatially:
```{r}
#bind the t object (chained) to the workers object. 
t <- t %>% filter(is.na(chained)) # only including origins
t <- t[,1:20] 
OD_v_HW_chained <- rbind(OD_v_HW, t)
```

```{r}
#importing zones
zn <- st_read("data-raw/zn.shp")
ALLLEZ <- st_read("data-raw/Zona_Bajas_Emisiones_MADRID.shp")

temp <- read_excel("data-raw/ZonificacionZT1259 complete.xlsx") %>% dplyr::select(ZT1259, Municipality)
zn_t <- zn %>% merge(temp, by = c("ZT1259"), all.x=TRUE)
```

```{r}
Otrips_zona_c <- OD_v_HW_chained %>% group_by(VORIZT1259)  %>% summarise(Otrips = (n())) # we group by the origin ZT1259 ID and then we summarize the number of rows that have the same ZT using the summarize() function and the n() function to create a new 'Otrips' column.

zn_c <- zn %>% merge(Otrips_zona_c, by.x = "ZT1259", by.y="VORIZT1259", all.x = TRUE)

@@

Otrips_zona <- OD_v_HW %>% group_by(VORIZT1259)  %>% summarise(Otrips = (n())) # we group by the origin ZT1259 ID and then we summarize the number of rows that have the same ZT using the summarize() function and the n() function to create a new 'Otrips' column.

zn_noc <- zn %>% merge(Otrips_zona, by.x = "ZT1259", by.y="VORIZT1259", all.x = TRUE)

@@

so as we can see... the number of origin ZNs with NO origin trips only goes down by 2.. (from 344 to 342). This is not a lot.

Next step... let's try to see what is happening in these NA zones. What sort of origin trips are being generated? 
```{r}
missing_workers <- zn_noc %>% filter(is.na(Otrips)) %>% dplyr::select(ZT1259) #344 zonas are missing home to work trips.

missing_workers <- missing_workers %>% merge(OD_v, by.x = "ZT1259", by.y="VORIZT1259", all.x = TRUE)

temp <- missing_workers %>% st_drop_geometry() %>% group_by(ZT1259) %>% summarise(trip_count = n(),
                                                                                  Owork_perc = sum(VORI == 2)/n(),
                                                                                  Ostudy_perc = sum(VORI == 4)/n(),
                                                                                  Osportwalk_perc = sum(VORI == 10)/n(),
                                                                                  Oshop_perc = sum(VORI == 5)/n(),
                                                                                  Ohome_perc = sum(VORI == 1)/n())
temp
```
```{r}
summary(temp)
```
```{r}
# write.csv(missing_workers %>% st_drop_geometry(),"ZT_Trips_missingworkers.csv", row.names = FALSE)
```

‘trip count’ is how many trips there are in each ZT. ‘Owork_perc’ is an origin work purpose (of that trip) in each ZT, i.e. 39.47% of origin trips are work related averaged across all ZTs. ‘Ostudy_perc’ is for study origin trips.. etc. You can see that the majority of origin trip purposes are work. This means that these are predominately ‘work’ related areas and none of the households surveyed live in those ZTs.

And looking in excel. even if we consider all origin zones from work themed chained trips (i.e., home to somewhere to work) as 'residential locations' (this is incorrect) we STILL have ~170 zones with no work-to-home trips.

@@

Then create the impedance functions:
```{r fitting-impedance-function}
# using fitdist function to fit a distribution using the default maximum likelihood estimation method and Nelder-Mead method for direct optimization
lnorm_ <- fitdistrplus::fitdist(data=transit_HW_alltt$TIEMPO_VIAJES,#%>% unlist() %>% as.numeric(),
                                "lnorm", 
                                method="mle",
                                #Notice the use of the weights,
                                weights = round(transit_HW_alltt$ELE_G_POND_ESC2),
                                optim.method="Nelder-Mead")
HW_transit_lnorm <- lnorm_

@@

lnorm_ <- fitdistrplus::fitdist(data=walk_HW_alltt$TIEMPO_VIAJES, #%>% unlist() %>% as.numeric(),
                                "lnorm", 
                                method="mle",
                                #Notice the use of the weights,
                                weights = round(walk_HW_alltt$ELE_G_POND_ESC2),
                                optim.method="Nelder-Mead")
HW_walk_lnorm <- lnorm_
```

@@

x_transit <- x %>%
  mutate(f = dlnorm(t,
                    HW_transit_lnorm$estimate[1],
                    HW_transit_lnorm$estimate[2]),
         mode = "transit")

@@

x_walk <- x %>%
  mutate(f = dlnorm(t,
                    HW_walk_lnorm$estimate[1],
                    HW_walk_lnorm$estimate[2]),
         mode = "walk")
```

@@

# The workers are classified by mode
workers <- OD_v_HW %>%
  group_by(VORIZT1259,
           mode) %>%
  # The sum is of the weighted workers
  summarize(workers = sum(ELE_G_POND_ESC2),
            .groups = "drop") %>%
  rename(ZT1259 = VORIZT1259)

@@

# Do we need to classify the jobs by mode?
jobs <- OD_v_HW %>%
  group_by(VDESZT1259) %>%
  summarize(jobs = sum(ELE_G_POND_ESC2),
            .groups = "drop") %>%
  rename(ZT1259 = VDESZT1259)
```

Next join workers and jobs to the ZN simple features object:
```{r}
load("data/zn.RData")

# zn <- zn %>%
#   # Join workers
#   left_join(workers,
#             by = "ZT1259") %>%
#   # Join jobs
#   left_join(jobs,
#             by = "ZT1259") %>%
#   # Replace NAs with zeros
#   mutate(workers_transit = tidyr::replace_na(workers_transit, 0),
#          workers_car = tidyr::replace_na(workers_car, 0),
#          workers_bike = tidyr::replace_na(workers_bike, 0),
#          workers_walk = tidyr::replace_na(workers_walk, 0),
#          jobs_transit = tidyr::replace_na(jobs_transit, 0),
#          jobs_car = tidyr::replace_na(jobs_car, 0),
#          jobs_bike = tidyr::replace_na(jobs_bike, 0),
#          jobs_walk = tidyr::replace_na(jobs_walk, 0))

workers_sf <- workers %>%
  # Join workers
  left_join(zn,
            by = "ZT1259") %>%
  st_as_sf()

jobs_sf <- jobs %>%
  # Join jobs
  left_join(zn,
            by = "ZT1259") %>%
  st_as_sf()

@@

OD <- OD_v_HW %>%
  group_by(VORIZT1259, 
           VDESZT1259,
           mode) %>%
  summarize(tt = mean(TIEMPO_VIAJES),
            .groups = "drop")

@@

Join workers and jobs to OD table:
```{r}
OD_2 <- OD %>%
  left_join(workers,
            by = c("VORIZT1259" = "ZT1259", "mode"))

OD_2 <- OD_2 %>%
  left_join(jobs,
            by = c("VDESZT1259" = "ZT1259"))
```

@@

OD_2 <- OD_2 %>%
  mutate(f = case_when(mode == "transit" ~ dlnorm(tt,
                                                  HW_transit_lnorm$estimate["meanlog"], 
                                                  HW_transit_lnorm$estimate["sdlog"]),
                       mode == "car/motor" ~ dgamma(tt,
                                                    HW_car_motor_gamma$estimate["shape"],
                                                    HW_car_motor_gamma$estimate["rate"]),
                       mode == "bike" ~ dlnorm(tt,
                                               HW_bike_lnorm$estimate["meanlog"],
                                               HW_bike_lnorm$estimate["sdlog"]),
                       mode == "walk" ~ dlnorm(tt, 
                                               HW_walk_lnorm$estimate["meanlog"], 
                                               HW_walk_lnorm$estimate["sdlog"])))
```

@@

Calculate spatial availabilty:
```{r}
OD_SA <- OD_2 %>%
  mutate(catch = 1) %>%
  mutate(SA_ij = sp_avail(., 
                          o_id = VORIZT1259,
                          d_id = VDESZT1259,
                          pop = workers,
                          opp = jobs,
                          r = catch,
                          f = f))
```

Check that the totals are maintained:
```{r}
sum(OD_SA$SA_ij)
sum(jobs$jobs)
```

@@

Spatial availability by origin (with check):
```{r}
SA_i <- OD_SA %>%
  group_by(VORIZT1259) %>%
  summarize(V_i = sum(SA_ij))

@@

Spatial availability by origin/mode (with check):
```{r}
SA_im <- OD_SA %>%
  group_by(VORIZT1259, 
           mode) %>%
  summarize(V_im = sum(SA_ij),
            .groups = "drop")

@@

Calculate worker population by zone of origin (with check):
```{r}
population <- workers %>%
  group_by(ZT1259) %>%
  summarize(population = sum(workers))

@@

Join zonal worker population (with checks):
```{r}
SA_i <- SA_i %>%
  left_join(population,
            by = c("VORIZT1259" = "ZT1259"))
sum(SA_i$V_i)

SA_im <- SA_im %>%
  left_join(population,
            by = c("VORIZT1259" = "ZT1259"))
sum(SA_im$V_im)
```

@@

Join geometry of zones to SA:
```{r}
SA_i <- SA_i %>%
  left_join(zn %>%
              dplyr::select(ZT1259),
            by = c("VORIZT1259" = "ZT1259")) %>%
  st_as_sf()

SA_im <- SA_im %>%
  left_join(zn %>%
              dplyr::select(ZT1259),
            by = c("VORIZT1259" = "ZT1259")) %>%
  st_as_sf()
```

Preliminaries to plotting:
```{r create-bounding-boxes}
#create the bounding box for the Toronto region + a little extra
bbox_new1 <- st_bbox(zn%>%filter(Municipality == "Madrid")) # current bounding box

@@

Madrid_muni_border <- zn %>%
  filter(Municipality == "Madrid") %>%
  st_union()
```

@@

Plot v (spatial availability per capita) by origin:
```{r}
ggplot() +
  geom_sf(data = SA_im,
          aes(fill = v_im),
          color = NA) +
  scale_fill_gradient2(name = expression(v["i"]),
                       midpoint = 1,
                       na.value = "grey90")+
  geom_sf(data = Madrid_muni_border, # border for Madrid muni
          colour=alpha("black",1), 
          size = 0.5, 
          fill=NA, 
          show.legend = "polygon") +
  annotation_north_arrow(location = "tl", # north arrow for both the main plot and inset
                         height = unit(0.8, "cm"), 
                         width = unit(0.8, "cm"),
                         style = north_arrow_orienteering(line_width = 0.25,
                                                          line_col = "dimgrey", 
                                                          fill = c("white","dimgrey"))) +
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm")) +
  theme_void() +
  theme(legend.position = c(1.10, 0.25), 
        plot.margin = margin(t = 0, r = 0, b = 0, l = -80)) + # positing legend and clipping out white space
  facet_wrap(~mode, nrow = 2)
```
<!--

@@

First, let's re-assign to-madrid muni car trips to transit trips:
```{r}
Madrid_muni <- zn %>% filter(Municipality == "Madrid") %>% 
  dplyr::transmute(ZT1259,
                   MM = 1)

OD_v_HW_optimistic <- OD_v_HW  %>% merge(Madrid_muni, by.x=("VDESZT1259"), by.y=("ZT1259"), all.x=T)

#reassigning car trips to transit
OD_v_HW_optimistic1 <- OD_v_HW_optimistic  %>% filter(MM == 1) %>% rowwise() %>%
  mutate(trips_transit = sum(trips_transit,trips_car_motor, na.rm = T),
         trips_car_motor = 0) #16627 ODs are making trips INTO muni, this is 59% of all OD pairs

OD_v_HW_optimistic <- rbind(OD_v_HW_optimistic1,OD_v_HW_optimistic2)

>>>>

Import the trip data:
```{r data-import-filter}
load(file = "data/OD_v.RData")
```

Keep only the direct home to work trips:
```{r}
# not chained home to work trips - direct
OD_v_HW <- OD_v %>% filter(VORI == "1" & VDES == "2") #only filter in origin casa trips and destination trabajo
```

Import zones:
```{r}
#importing zones
load("data/zn1259.RData")
load("data/zn208.RData")
load("data/zn_u.RData") # a key to join the zt1259 to the zn208 + contains municipality names
```

@@

This needs to be re-thought for the ZT208 resolution. Since we don't have an issue with missing data from ZT208 zones, we will skip trying to include chained trips (including also complicates the impedance zone):
<!-- Checking the trip data - it appears each OD is a row, i.e., chained trips are multiple rows. We want to extract every home origin (VORI = 1) to every work destination (VDEST =2) trip as a single row... so let's do this in multiple steps.

Next, let's identify the chained trips and then consolidate them. We only care about trip O/D, travel mode, travel time, and 
```{r}
temp <- OD_v %>% group_by(ID_HOGAR,ID_IND) %>% 
  filter((VORI == "1" & VDES != "2" & VDES != "1") | (VDES == "2" & VORI != "1"& VORI != "2")) %>% #here the origin NEEDS to be from home, the destination can be anywhere OR the origin can be from anywhere but the destination NEEDS to be work. 
  mutate(chained = lag(VDES) - VORI)

chained2 <- (ifelse(temp$chained[-1]==0 | temp$chained==0, 0, NA)) %>% as.data.frame() #identifying chained trips using '0's

temp <- cbind(temp,chained2) %>% dplyr::rename("chained2" = ".")

temp <- temp %>% filter(chained2==0) # filtering only CHAINED trips - reminder these consider only 1 location before work. 1,629 trips 
```

Now... let's only include chained trips with the same mode! 
```{r}
t <- temp %>% group_by(ID_HOGAR, ID_IND) %>% mutate(trip = ifelse(((ID_VIAJE - lag(ID_VIAJE)) == 1)
                                                               & ((MODO_PRIORITARIO - lag(MODO_PRIORITARIO))== 0), 
                                                               MODO_PRIORITARIO, 0)) #if trip is chained the mode (number) is printed, if the trip is not chained a '0' is printed

t %>% filter(trip == 0) %>% nrow() / t  %>% filter(chained == 0) %>% nrow()    #of chained trips / number of non-chained trips
t  %>% filter(trip != 0 & !is.na(trip))%>% nrow() / t %>% filter(chained == 0) %>% nrow() #of chained trips / number of non-chained trips
```
so ~4% of trips are chained, and ~27% of chained trips are not made using the same mode. THe rest of them are (`r temp %>% filter(trip != 0 & !is.na(trip)) %>% nrow() ` trips are chained (1 intermediate destination) and use the same mode).

@@

Let's see if this sample increases the number of ZTs spatially:
```{r}
#bind the t object (chained) to the workers object. 
t <- t %>% filter(is.na(chained)) # only including origins
t <- t[,1:24] 
OD_v_HW_chained <- rbind(OD_v_HW, t)
```

```{r}
Otrips_zona_c <- OD_v_HW_chained %>% group_by(VORI_CD_ZT208)  %>% summarise(Otrips = (n())) # we group by the origin ZT208 ID and then we summarize the number of rows that have the same ZT using the summarize() function and the n() function to create a new 'Otrips' column.

zn_c <- zn208 %>% merge(Otrips_zona_c, by.x = "CD_ZT208", by.y="VORI_CD_ZT208", all.x = TRUE)

@@

Otrips_zona <- OD_v_HW %>% group_by(VORI_CD_ZT208)  %>% summarise(Otrips = (n())) # we group by the origin ZT208 ID and then we summarize the number of rows that have the same ZT using the summarize() function and the n() function to create a new 'Otrips' column.

zn_noc <- zn208 %>% merge(Otrips_zona, by.x = "CD_ZT208", by.y="VORI_CD_ZT208", all.x = TRUE)

@@

so as we can see... the number of origin ZNs with NO origin trips doesn't go down - we only have 1 missing ZT using the 208 zonig system and 1 is still missing after including the chained trips.

Next step... let's try to see what is happening in these NA zones. What sort of origin trips are being generated? (analysis for the 1259 zoning system)-->

<!-- ```{r} -->
<!-- missing_workers <- zn_noc %>% filter(is.na(Otrips)) %>% dplyr::select(ZT1259) #344 zonas are missing home to work trips. -->

<!-- missing_workers <- missing_workers %>% merge(OD_v, by.x = "ZT1259", by.y="VORIZT1259", all.x = TRUE) -->

<!-- temp <- missing_workers %>% st_drop_geometry() %>% group_by(ZT1259) %>% summarise(trip_count = n(), -->
<!--                                                                                   Owork_perc = sum(VORI == 2)/n(), -->
<!--                                                                                   Ostudy_perc = sum(VORI == 4)/n(), -->
<!--                                                                                   Osportwalk_perc = sum(VORI == 10)/n(), -->
<!--                                                                                   Oshop_perc = sum(VORI == 5)/n(), -->
<!--                                                                                   Ohome_perc = sum(VORI == 1)/n()) -->
<!-- temp -->
<!-- ``` -->
<!-- ```{r} -->
<!-- summary(temp) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- # write.csv(missing_workers %>% st_drop_geometry(),"ZT_Trips_missingworkers.csv", row.names = FALSE) -->
<!-- ``` -->

<!-- ‘trip count’ is how many trips there are in each ZT. ‘Owork_perc’ is an origin work purpose (of that trip) in each ZT, i.e. 39.47% of origin trips are work related averaged across all ZTs. ‘Ostudy_perc’ is for study origin trips.. etc. You can see that the majority of origin trip purposes are work. This means that these are predominately ‘work’ related areas and none of the households surveyed live in those ZTs.  -->

<!-- And looking in excel. even if we consider all origin zones from work themed chained trips (i.e., home to somewhere to work) as 'residential locations' (this is incorrect) we STILL have ~170 zones with no work-to-home trips.  -->

@@

Then create the impedance functions:
```{r fitting-impedance-function}
# using fitdist function to fit a distribution using the default maximum likelihood estimation method and Nelder-Mead method for direct optimization
lnorm_ <- fitdistrplus::fitdist(data=transit_HW_alltt$TIEMPO_VIAJES,#%>% unlist() %>% as.numeric(),
                                "gamma", 
                                method="mle",
                                #Notice the use of the weights,
                                weights = round(transit_HW_alltt$ELE_G_POND_ESC2),
                                optim.method="Nelder-Mead")
HW_transit_lnorm <- lnorm_

@@

logis_ <- fitdistrplus::fitdist(data=walk_HW_alltt$TIEMPO_VIAJES, #%>% unlist() %>% as.numeric(),
                                "logis", 
                                method="mle",
                                #Notice the use of the weights,
                                weights = round(walk_HW_alltt$ELE_G_POND_ESC2),
                                optim.method="Nelder-Mead")
HW_walk_lnorm <- logis_
```

@@

x_transit <- x %>%
  mutate(f = dgamma(t,
                    HW_transit_lnorm$estimate[1],
                    HW_transit_lnorm$estimate[2]),
         mode = "transit")

@@

x_walk <- x %>%
  mutate(f = dlogis(t,
                    HW_walk_lnorm$estimate[1],
                    HW_walk_lnorm$estimate[2]),
         mode = "walk")
```

@@

# The workers are classified by mode
workers <- OD_v_HW %>%
  group_by(VORI_CD_ZT208,
           mode) %>%
  # The sum is of the weighted workers
  summarize(workers = sum(ELE_G_POND_ESC2),
            .groups = "drop") %>%
  rename(CD_ZT208 = VORI_CD_ZT208)

@@

# Do we need to classify the jobs by mode?
jobs <- OD_v_HW %>%
  group_by(VDES_CD_ZT208) %>%
  summarize(jobs = sum(ELE_G_POND_ESC2),
            .groups = "drop") %>%
  rename(CD_ZT208 = VDES_CD_ZT208)
```

Next join workers and jobs to the ZN simple features object:
```{r}
# zn <- zn %>%
#   # Join workers
#   left_join(workers,
#             by = "ZT1259") %>%
#   # Join jobs
#   left_join(jobs,
#             by = "ZT1259") %>%
#   # Replace NAs with zeros
#   mutate(workers_transit = tidyr::replace_na(workers_transit, 0),
#          workers_car = tidyr::replace_na(workers_car, 0),
#          workers_bike = tidyr::replace_na(workers_bike, 0),
#          workers_walk = tidyr::replace_na(workers_walk, 0),
#          jobs_transit = tidyr::replace_na(jobs_transit, 0),
#          jobs_car = tidyr::replace_na(jobs_car, 0),
#          jobs_bike = tidyr::replace_na(jobs_bike, 0),
#          jobs_walk = tidyr::replace_na(jobs_walk, 0))

workers_sf <- workers %>%
  # Join workers
  left_join(zn208,
            by = "CD_ZT208") %>%
  st_as_sf()

jobs_sf <- jobs %>%
  # Join jobs
  left_join(zn208,
            by = "CD_ZT208") %>%
  st_as_sf()

@@

OD <- OD_v_HW %>%
  group_by(VORI_CD_ZT208,
           VDES_CD_ZT208,
           mode) %>%
  summarize(tt = mean(TIEMPO_VIAJES),
            .groups = "drop")

@@

Join workers and jobs to OD table:
```{r}
OD_2 <- OD %>%
  left_join(workers,
            by = c("VORI_CD_ZT208" = "CD_ZT208", "mode"))

OD_2 <- OD_2 %>%
  left_join(jobs,
            by = c("VDES_CD_ZT208" = "CD_ZT208"))
```

@@

OD_2 <- OD_2 %>%
  mutate(f = case_when(mode == "transit" ~ dgamma(tt,
                                                  HW_transit_lnorm$estimate["shape"], 
                                                  HW_transit_lnorm$estimate["rate"]),
                       mode == "car/motor" ~ dgamma(tt,
                                                    HW_car_motor_gamma$estimate["shape"],
                                                    HW_car_motor_gamma$estimate["rate"]),
                       mode == "bike" ~ dlnorm(tt,
                                               HW_bike_lnorm$estimate["meanlog"],
                                               HW_bike_lnorm$estimate["sdlog"]),
                       mode == "walk" ~ dlogis(tt, 
                                               HW_walk_lnorm$estimate["location"], 
                                               HW_walk_lnorm$estimate["scale"])))
```

@@

Calculate spatial availabilty:
```{r}
OD_SA <- OD_2 %>%
  mutate(catch = 1) %>%
  mutate(SA_ij = sp_avail(., 
                          o_id = VORI_CD_ZT208,
                          d_id = VDES_CD_ZT208,
                          pop = workers,
                          opp = jobs,
                          r = catch,
                          f = f))
```

Check that the totals are maintained:
```{r}
sum(OD_SA$SA_ij, na.rm = T)
sum(jobs$jobs)
```

@@

Spatial availability by origin (with check):
```{r}
SA_i <- OD_SA %>%
  group_by(VORI_CD_ZT208) %>%
  summarize(V_i = sum(SA_ij))

@@

Spatial availability by origin/mode (with check):
```{r}
SA_im <- OD_SA %>%
  group_by(VORI_CD_ZT208, 
           mode) %>%
  summarize(V_im = sum(SA_ij),
            .groups = "drop")

@@

Calculate worker population by zone of origin (with check):
```{r}
population <- workers %>%
  group_by(CD_ZT208) %>%
  summarize(population = sum(workers))

@@

Join zonal worker population (with checks):
```{r}
SA_i <- SA_i %>%
  left_join(population,
            by = c("VORI_CD_ZT208" = "CD_ZT208"))
sum(SA_i$V_i)

SA_im <- SA_im %>%
  left_join(population,
            by = c("VORI_CD_ZT208" = "CD_ZT208"))
sum(SA_im$V_im)
```

@@

Join geometry of zones to SA:
```{r}
SA_i <- SA_i %>%
  left_join(zn208 %>%
              dplyr::select(CD_ZT208),
            by =c("VORI_CD_ZT208" = "CD_ZT208")) %>%
  st_as_sf()

SA_im <- SA_im %>%
  left_join(zn208 %>%
              dplyr::select(CD_ZT208),
            by = c("VORI_CD_ZT208" = "CD_ZT208")) %>%
  st_as_sf()
```

Preliminaries to plotting:

Figure out which CD_ZT208 correspond to Madrid - we see that CD_ZT208 1 through 59 do so.

```{r}
t <- zn_u %>% dplyr::select(CD_ZT208, Municipality) %>% group_by(CD_ZT208,Municipality) %>% summarize() %>% filter(Municipality == "Madrid")
t
```

```{r create-bounding-boxes}
zn208 <- zn208 %>% mutate(MM = ifelse(str_starts(ZT208,"Madrid"), 1, 0)) %>% filter(MM == 1)

#create the bounding box for the Toronto region + a little extra
bbox_new1 <- st_bbox(zn208 %>%filter(MM == 1)) # current bounding box

@@

sf::sf_use_s2(FALSE) #this terns off the spherical geometry component in the st_union() function - this assumes the sf is just planar. Thats' fine for our purposes
Madrid_muni_border <- zn208 %>%
  filter(MM == 1) %>%
  st_union()
```

@@

Plot v (spatial availability per capita) by origin:
```{r}
ggplot() +
  geom_sf(data = SA_im,
          aes(fill = v_im),
          color = NA) +
  scale_fill_gradient2(name = expression(v["i"]),
                       midpoint = 1,
                       na.value = "grey90")+
  geom_sf(data = Madrid_muni_border, # border for Madrid muni
          colour=alpha("black",1), 
          size = 0.5, 
          fill=NA, 
          show.legend = "polygon") +
  annotation_north_arrow(location = "tl", # north arrow for both the main plot and inset
                         height = unit(0.8, "cm"), 
                         width = unit(0.8, "cm"),
                         style = north_arrow_orienteering(line_width = 0.25,
                                                          line_col = "dimgrey", 
                                                          fill = c("white","dimgrey"))) +
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm")) +
  theme_void() +
  theme(legend.position = c(1.10, 0.25), 
        plot.margin = margin(t = 0, r = 0, b = 0, l = -80)) + # positing legend and clipping out white space
  facet_wrap(~mode, nrow = 2)
```

The previous attempts at calculating spatial availability is commented out and included in the end of the document. There were three different approaches, and the approach reflective above is the 3rd approach tested. The first approach can be used in the case of time windows, but the 2nd and 3rd are similar just the 3rd approach is more conceptually easy to understand and code.

<!-- AS finished here May 31, am stuck on how to shift car trips to transit (in a way that's flexible and will allow us to easily toggle the modal shift scenarios (i.e., instead of 100% transit, maybe 80%/10/10 transit/bike/walk)) -->

First, let's re-assign to-madrid muni car trips to transit trips:
```{r}
Madrid_muni <- zn208 %>% filter(MM == 1)

# OD <- OD_v_HW %>%
#   group_by(VORI_CD_ZT208,
#            VDES_CD_ZT208,
#            mode) %>%
#   summarize(tt = mean(TIEMPO_VIAJES),
#             .groups = "drop")

OD_v_HW_VDES_Mad <- OD_v_HW %>% filter(VDES_Municipality == "Madrid") %>%
  group_by(VORI_CD_ZT208, VDES_CD_ZT208, mode) %>% 
  summarize(TIEMPO_VIAJES = mean(TIEMPO_VIAJES))

OD_v_HW_VDES_Mad %>% mutate(tt = ifelse(mode == "car", ))

OD_v_HW_optimistic1 %>% mutate(TIEMPO_VIAJES_opt = ifelse(mode="car" & MM == 1, NA, ))

OD_v_HW_optimistic1 %>% group_by(mode) %>% summarize(car = mean())

  mutate(TIEMPO_VIAJES_opt = ifelse(mode = "car", ))
```

```{r}
#reassigning car trips to transit trips (i.e., a 100% shift to transit from car)
OD_v_HW_optimistic1 <- OD_v_HW_optimistic  %>% filter(MM == 1) %>% #only CAR trips to Madrid
  mutate(trips_transit = sum(trips_transit,trips_car_motor, na.rm = T), #only trips 
         trips_car_motor = 0) #16627 ODs are making trips INTO muni, this is 59% of all OD pairs

OD_v_HW_optimistic <- rbind(OD_v_HW_optimistic1,OD_v_HW_optimistic2,OD_v_HW_optimistic2)

The commented out different spatial availability calculation approaches (also done at the ZT1259 resolution instead of ZT208)
<!--

@@
